---
title: "SDaouki_Week5_discussion"
author: "Saloua Daouki"
date: "2024-02-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CoinTosses Program 1

This is just a simple coin toss function that prints the number of heads and the number of tails come up depends on the number $n$:

```{r number of H & T}
# Function to simulate coin tosses
CoinToss1 <- function(n) {
  # Generate random coin toss results (0 for tails, 1 for heads)
  toss_results <- sample(0:1, n, replace = TRUE)
  
  # Count the number of heads and tails
  n_heads <- sum(toss_results == 1)
  n_tails <- sum(toss_results == 0)
  
  # Return the results
  return(list(heads = n_heads, tails = n_tails))
}

# Excecute the function:
n <- 1000
results <- CoinToss1(n)
cat("Number of heads:", results$heads, "\n")
cat("Number of tails:", results$tails, "\n")

```

This function returns a series of H & T depends on the number $n$;

```{r H & T series}
# Function to simulate coin tosses and return a series of 'H' and 'T'
CoinToss11 <- function(n1) {
  # Generate random coin toss results (0 for tails, 1 for heads)
  toss_results <- sample(0:1, n1, replace = TRUE)
  
  # Convert numeric results to characters ('H' for heads, 'T' for tails)
  toss_series <- ifelse(toss_results == 1, 'H', 'T')
  
  # Return the series of 'H' and 'T'
  return(toss_series)
}

# Excecute the function:
n1 <- 10
toss_series <- CoinToss11(n1)
cat("Coin toss series:", paste(toss_series, collapse = ' '), "\n")
```

## CointToss Program 2

```{r}
# Function to simulate coin tosses and return the proportion of heads minus 1/2 for the entire set
CoinToss2 <- function(n2) {
  # Initialize variables to count heads and tails
  n_H <- 0  # Number of heads obtained
  n_T <- 0  # Number of tails obtained
  
  # Loop through each toss
  for (i in 1:n2) {
    # Generate random coin toss results (0 for tails, 1 for heads)
    toss_result <- sample(0:1, 1)
    
    # Update counts
    n_H <- n_H + toss_result
    n_T <- n_T + (1 - toss_result)  # Count tails as well
  }
  
  # Calculate proportion of heads minus 1/2 for the entire set
  proportion <- (n_H / n2) - 0.5
  
  # Return the proportion
  return(proportion)
}

# Excecute the function:
n2 <- 150000
Result <- CoinToss2(n2)
cat("Proportion of heads minus 1/2 for the entire set:", Result, "\n")

```

I tried more than 10 number for the number of tosses and organized the data into the following table that I converted to a data frame in order to be able to plot the results and visualize the relationship; 

```{r dataframe}
# Create a list of values
values_list <- list(
  n2 = c(200, 300, 800, 1000, 1500, 2000, 5000, 10000, 50000, 100000, 150000),
  First100Proportion = c(0.04,0.01667,-0.0025,0.004,-0.00667,0.014,-0.0192, -0.0024, 0.00268, 0.00236, 0.00102)
)

# Convert the list to a data frame
values_df <- data.frame(values_list)

# Print the data frame
print(values_df)

```


```{r}
library(ggplot2)
```

```{r Plot 1 for the first 100 tosses}
# Create a ggplot 
Proportion1 <- ggplot(data = values_df, aes(x = n2, y = First100Proportion)) +
  geom_line(color = "blue") +  # Add a line plot with blue color
  labs(x = "number of tosses", y = "Proportion of heads minus 1/2", title = "Proportions Vs number of tosses")  # Add axis labels and title

# Display the plot
print(Proportion1)

```

Based on the table and the line graph above, when the number of tosses is less than 5000, the proportion ranges between 0.04 and -0.00240. In other words, there is no clear pattern whether the proportion getting close to zero or not as the the number of tosses increases. However, when the number of tosses increases after 5000, the proportion gets closer to zero.


## CoinToss Program 3


```{r}
# Function to simulate coin tosses and return the proportion of heads minus 1/2 and the number of heads minus half the number of tosses for the entire set
CoinToss3 <- function(n3) {
  # Initialize variables to count heads and tails
  n_H3 <- 0  # Number of heads obtained
  n_T3 <- 0  # Number of tails obtained
  
  # Loop through each toss
  for (i in 1:n3) {
    # Generate random coin toss results (0 for tails, 1 for heads)
    toss_result3 <- sample(0:1, 1)
    
    # Update counts
    n_H3 <- n_H3 + toss_result3
    n_T3 <- n_T3 + (1 - toss_result3)  # Count tails as well
  }
  
  # Calculate both proportion of heads minus 1/2 and the number of heads minus half the number of tosses
  proportionH <- (n_H3 / n3) - 0.5
  numberH <- n_H3 - (n3/2)
  
  # Return both values as a list or a data frame
  return(list(proportion = proportionH, number_heads = numberH))
}

# Example usage:
n3 <- 150000
Result3 <- CoinToss3(n3)
cat("Proportion of heads minus 1/2 for the entire set:", Result3$proportion, "\n")
cat("Number of heads minus half the number of tosses:", Result3$number_heads, "\n")

```



```{r dataframe1}
# Create a list of values
values_list1 <- list(
  n3 = c(200, 300, 800, 1000, 1500, 2000, 5000, 10000, 50000, 100000, 150000),
  Proportion3 = c(0.055,0.01667,0.0075, 0.003, 0.01333, 0.0075, 0.0028, -0.002, 0.00232, -0.0005, -0.0002 ),
  NumberOfH = c(11, 5, 6, 3, 20, 15, 14, -20, 116, -50 , -30)
)

# Convert the list to a data frame
values_df1 <- data.frame(values_list1)

# Print the data frame
print(values_df1)

```



```{r Plot 2 for Proportion3}
# Create a ggplot 
Plot2 <- ggplot(data = values_df1, aes(x = n3, y = Proportion3)) +
  geom_line(color = "darkgreen") +  # Add a line plot with dark green color
  labs(x = "number of tosses", y = "Proportion of heads minus 1/2", title = "Proportions Vs number of tosses")  # Add axis labels and title

# Display the plot
print(Plot2)

```

The proportion of heads minus 1/2 gets closer to zero as n increases; we can see it more clearly from the line plot especially when the number of tosses is greater than 2000.

Now let's plot the number of heads minus 1/2 the number of tosses vs the number of tosses;

```{r Plot 3 for Number of heads minus half number of tosses}
# Create a ggplot 
Plot3 <- ggplot(data = values_df1, aes(x = n3, y = NumberOfH)) +
  geom_line(color = "magenta") +  # Add a line plot with dark green color
  labs(x = "number of tosses", y = "Number of heads minus 1/2 number of tosses", title = "H-(n/2) Vs number of tosses")  # Add axis labels and title

# Display the plot
print(Plot3)

```
Based on the table and the graph, it seems that there is a peak in the number of heads minus 1/2 the number of tosses (numberOfH) when the number of tosses is 5000, after that, the numberOfH decreases drastically from 116 to -50. The only smooth decrease in the numberOfH was in the ranges [0,1000] and [1500, 10000].

