---
title: "SDaouki_Assignment7"
author: "Saloua Daouki"
date: "2024-03-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1:

Let X1, X2, . . . , Xn be $n$ ***mutually independent random variables***, each of which is uniformly distributed on the integers from 1 to k. **Let Y denote the minimum of the Xi’s**. Find the distribution of Y.

Step1: The probability of $X_i$ to be greater than a value, let's say $l$, where $1 \leq l < k$ will be the number of outcomes greater than $l$ divided by the total number of outcomes which is $k$. Since there are $n$ independent random variables, then:

$P(X_i > l) = (\frac {k-l}{k})^n$ (1)

Now, to find the probability that the minimum of all $X_i$s (denoted by Y) is equal to $l$ is:

$P(Y=l) = P(Y>l-1) - P(Y>l)$ (2)

Using the equation (1), we can substitute into equation (2):

$P(Y=l) = \left(\frac{k-(l-1)}{k} \right)^n - \left(\frac{k-l}{k} \right)^n$

$\Rightarrow P(Y=1) = \left(\frac{k-l+1}{k} \right)^n - \left(\frac{k-l}{k} \right)^n$

## Exercise 2:

Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).

a.  What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. ***Model as a geometric.*** (Hint: the probability is equivalent to not failing during the first 8 years..)

$\Rightarrow$ **The solution**:

The probability that the MRI will fail in any year during the 10 years is $p=\frac{1}{10}$, then the probability of not failing will be $q=1-p = 1-\frac{1}{10}=\frac{9}{10}$

So, $P(\text{not failing during the 8 years}) = q^8 \times p = (\frac{9}{10})^8 \times \frac{1}{10}$

```{r}
# Define the probability of failure in a given year
p <- 1 / 10
# Probability of not failing 
q <- 9 / 10

# Probability of not failing in 8 years
P_fail_in_8 <- (q^8) * p
cat("Probability of not failing during the 8 years is :", P_fail_in_8, "\n")
```

The expected value (the mean) will be just 10 years, since that is the lifetime of the MRI machine.

The standard deviation is $\sigma = \frac {\sqrt{1-p}}{p}$

```{r}
SD_value <- sqrt((1 - p) / p^2)
cat("Standard deviation:", SD_value, "years\n")
```


b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. ***Model as an exponential.***

$\Rightarrow$ **The solution**:

The probability that the MRI will fail after 8 years (before its lifetime) is $1-P(\text{not failing during the 8 years})$

$P(\text{not failing during the 8 years}) = e^{-\lambda \cdot 8}$ where $\lambda = 1/10$ 

```{r}
eight_year_fail <- 1- exp((-1/10)*8)
cat("The probability that the MRI will fail after 8 years is", eight_year_fail)
```

The expected failure time is the expected value; 10 years.

Since this is an exponential distribution, the standard deviation is the same as the mean; so $SD=10$

c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. ***Model as a binomial.** *(Hint: 0 success in 8 years)

$\Rightarrow$ **The solution**:

$\Rightarrow$ The binomial formula is: $P(X=k)=\binom{n}{k} p^k (1-p)^{n-k}$
Where: $n=8$, $ k=0$ and given the rate of the failure is only once per 10 years, then the $P(\text{failure}) =0.1$ while $P(\text{success}) = 0.9$

Then; $P(\text{failing after 8 years}) = \binom{8}{0} p^0 (1-p)^{8} = \binom{8}{0} (0.9)^0 (0.1)^{8} $

And:  the expected value is: $E=n \times p = 8 \times 0.9$

  The standard deviation is: $\sigma =\sqrt {n \times p \times (1-p)}$
  
```{r}
# Defining the coefficients:
n <- 8
k <- 0
p <- 0.9

# The probability of failing after 8 years
Fail_after_8years <- (p^0 ) * ((1-p)^8)

# Expected Value
E_Value <- n * p

# Standard deviation
StandardV <- (n*p*(1-p))^(1/2)

# Display all values calculated above
cat ("The probability that the machine will fail after 8 years is ", Fail_after_8years, "\n", "The Expected Value is ", E_Value, "years", "\n", "The standard Deviation is ", StandardV, "years", "\n")
```
 
d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. ***Model as a Poisson. ***

$\Rightarrow$ **The solution**: 

The Poisson formula is $P(X=k) = \frac{\lambda^k}{k!} e^{-\lambda}$

So, the probability that the machine will fail after 8 years is:
 $P(\text{failing after 8 years} = 1 - P(\text{not failing in 8 years}))$
 And $P(\text{not failing in 8 years} = \sum_{k=0}^{8} \frac {\lambda^k}{k!} e^{-\lambda}$
 
The Expected value is $E(X) = \lambda$ and the standard deviation is $\sigma = \sqrt {\lambda}$

```{r}
# Define average
lambda <- 8

# Calculate the probability of not failing
for (i in 0:8) {
    prob <- (lambda^i / factorial(i)) * exp(-lambda)
}
# Probability of failing
prob_fail <- 1-prob
cat("Probability of machine failing after 8 years:", prob_fail, "\n")

# Expected value and standard deviation
ExpectedV <- lambda
SD_Value1 <- sqrt(lambda)
cat ("The expected value is ", ExpectedV, "\n" , "The standard Deiviation is ", SD_Value1, "\n")
```
      
      
      
      