---
title: "SDaouki_Assignment_8"
author: "Saloua Daouki"
date: "2024-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 11 page 303

***11. A company buys 100 lightbulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out?***

#### Given:

$n = 100$ lighbulbs

mean = 1000

#### So

$\lambda = \frac{1}{\text{mean}} = \frac{1}{1000}$

And since each of the $100$ lighbulbs has an **exponential lifetime** of $1000$ hours, then the expected time for the first of them to burn out is: (similar to example 6.29 page 275 in the probability text)

$E(\text{Time for the first lightbulb to burn out}) = \frac{1}{100 \lambda} = \frac{1}{100 \cdot \frac{1}{1000}} = \frac{1000}{100} = 10 $ hours.

## Exercise 14 Page 303:

***14. Assume that $X_1$ and $X_2$ are independent random variables, each having an exponential density with parameter $\lambda$. Show that $Z = X_1 − X_2$ has density $fZ(z) = \frac{1}{2} \lambda e^{-\lambda |z|}$***

The convolution theorem is: $fZ(z) = \int_{-\infty}^{\infty} fX_1(x) \cdot fX_2(x+z) dx$

We know that the density function of each of the random variable is:

$fX_1(x_1) = \lambda e^{- \lambda x_1}$

$fX_2 (x_2) = \lambda e^{-\lambda x_2}$


Substituting those densities in the convolution theorem:


$$
\begin{align*} 
f_Z(z) &= \int_{-\infty}^{\infty} \lambda e^{-\lambda x} \cdot \lambda e^{-\lambda (x+z)} dx \\
f_Z(z) &= \lambda ^2 \int_{-\infty}^{\infty} e^{-\lambda x} \cdot e^{-\lambda (x+z)} dx \\
f_Z(z) &= \lambda ^2 \int_{-\infty} ^{\infty} e^{- \lambda (2x+z)} dx \\
f_Z(z) &= \lambda ^2 e^{-\lambda z} \int_{- \infty} ^{\infty} e^{-2 \lambda x} dx \\
f_Z(z) &= \lambda ^2 e^{-\lambda z} [-\frac{1}{2 \lambda}e^{-2 \lambda x}]_{- \infty} ^{\infty} \\
f_Z(z) &= \frac{\lambda ^2}{2 \lambda} e^{-\lambda z} [e^{-\infty} - e^{\infty}] \\

f_Z(z) &= \frac{\lambda}{2} e^{-\lambda z} (0-0) \\
f_Z(z) &= \frac{\lambda}{2} e^{-\lambda z} \\
f_Z(z) &= \frac{1}{2} \lambda e^{- \lambda |z|}
\end{align*}
$$

Since $e^{-\lambda z}$ doesn't have $x$ and the integration with respect to x, then $e^{-\lambda z}$ is considered as a constant, so we can take it out of the integral.


## Exercise 1 Page 320-321

***Let $X$ be a continuous random variable with mean $\mu = 10$ and variance $\sigma ^2 = 100/3$. Using Chebyshev’s Inequality, find an upper bound for the following probabilities.***

(a) $P(|X−10|≥2)$.

(b) $P(|X−10|≥5)$. 

(c) $P(|X−10|≥9)$.

(d) $P(|X − 10| ≥ 20)$.


Chebyshev's Inequality says (page 305):

Let $X$ be a discrete random variable with expected value $\mu = E(X)$, and let $\epsilon > 0$ be any positive real number. Then

$P(|X- \mu | ) \geq  \epsilon ) \leq  \frac{V(X)}{\epsilon ^2}$

So the upper bound that we are looking for is $\frac{V(X)}{\epsilon ^2}$

And we can find the variance using: $V(X) = E((X - \mu)^2)$

But in this case $V(X)$ is given, which is $V(X)= \sigma ^2 = \frac{100}{3}$

#### (a) $P(|X−10|≥2)$.

Using the example 8.1 page 306 from the probability text,
The upper bound will be $\frac{1}{k^2}$ where in this case there 2 standard deviations. And $\epsilon = k \sigma$ $\Rightarrow k = \frac{\epsilon}{\sigma} = \frac{2}{\sqrt{\frac{100}{3}}} = \frac{2 \sqrt{3}}{10}$

```{r}
#Setting the parameters:
k <- (2* sqrt(3)) / 10

#calculate the upper bound
UpperBound_2 <- format(1/ k^2 , digits=3)

#Displaying the upper bound
cat("The upper bound of the probability with 2 standard deviations using Chebyshev's Inequality is " , UpperBound_2, "\n")
```


#### (b) $P(|X−10|≥5)$. 

In this case there are 5 standard deviations. And $\epsilon = k \sigma$ $\Rightarrow k = \frac{\epsilon}{\sigma} = \frac{5}{\sqrt{\frac{100}{3}}} = \frac{5 \sqrt{3}}{10}$

```{r}
#Setting the parameters:
k <- (5* sqrt(3)) / 10

#calculate the upper bound
UpperBound_5 <- format(1/ k^2 , digits=3)

#Displaying the upper bound
cat("The upper bound of the probability with 5 standard deviations using Chebyshev's Inequality is " , UpperBound_5, "\n")
```

#### (c) $P(|X−10|≥9)$. 

In this case there are 9 standard deviations. And $\epsilon = k \sigma$ $\Rightarrow k = \frac{\epsilon}{\sigma} = \frac{9}{\sqrt{\frac{100}{3}}} = \frac{9 \sqrt{3}}{10}$

```{r}
#Setting the parameters:
k <- (9* sqrt(3)) / 10

#calculate the upper bound
UpperBound_9 <- format(1/ k^2 , digits=3)

#Displaying the upper bound
cat("The upper bound of the probability with 9 standard deviations using Chebyshev's Inequality is " , UpperBound_9, "\n")
```

#### (d) $P(|X − 10| ≥ 20)$.

In this case there are 20 standard deviations. And $\epsilon = k \sigma$ $\Rightarrow k = \frac{\epsilon}{\sigma} = \frac{20}{\sqrt{\frac{100}{3}}} = \frac{20 \sqrt{3}}{10}$

```{r}
#Setting the parameters:
k <- (20* sqrt(3)) / 10

#calculate the upper bound
UpperBound_20 <- format(1/ k^2 , digits=3)

#Displaying the upper bound
cat("The upper bound of the probability with 20 standard deviations using Chebyshev's Inequality is " , UpperBound_20, "\n")
```